---
title: "My Project"
format: 
  gfm:
    preview-mode: raw
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE
)
```

# Bachelor Thesis

## Project Structure

`R/`: contains all R-scripts for data analysis and visualization

```{r}
fs::dir_tree("R/")
```

The `R/00_functions.R` files contains all custom R-functions that are defined for the project. The file h_*(1-4).R containts the analysis of the four hypotheses. The `R/combine_plots.R` file combines all plots generated for each hypothesis into one large multi-panel figure. The `R/07_check_validity.R` script containts diagnostic code to assess the problems within the parameter estimation of SPEEC. The `R/08_check_convergence.R` file contains a shiny app to assess the parameter convergence for each study. Finally `R/09_dispersion.R` contains code to analyse the overall dispersion of the estimated publication bias parameter.

`data/`: Data files

```{r}
fs::dir_tree("data/", type = "directory")
```

The data is generally structured into three subdirectories (meta, optim, src). The `meta` directory contains all data regarding the raw and processed meta-analytical data. The `optim` directory contains all data related to the parameter optimization (raw, and processed). Finally, the `src` folder contains additional data that is needed within the Quarto file to compile the thesis.

`scripts/`: Quarto scripts that are compiled into a reproducible report

```{r}
fs::dir_tree("scripts/")
```

## Compilation

To automatically compile the thesis, run the following bash script in the terminal:

``` bash
./build.sh
```