```{=tex}
\newpage
\pagenumbering{roman}
\thispagestyle{empty}
\newpage
\mbox{}\thispagestyle{empty}
\newpage
```

# Abstract {.unnumbered}

```{=tex}
\newpage
\begingroup
\setstretch{1.3}
\setcounter{tocdepth}{2}
\tableofcontents
\endgroup
\newpage
\pagenumbering{arabic}

```


```{r}
#| echo: false
#| label: setup

knitr::opts_chunk$set(
  echo = FALSE,
  results = "asis",
  fig.align = "center",
  message = FALSE
)

library(here)

```

# Introduction

```{r}
#| echo: false
#| results: "asis"
#| label: introduction

intro <- knitr::knit_child(here("scripts/1_introduction.qmd"), quiet = TRUE)
cat(intro, sep = "\n")
```


```{r}
#| echo: false
#| results: "asis"
#| label: methods

method <- knitr::knit_child(here("scripts/2_methods.qmd"), quiet = TRUE)
cat(method, sep = "\n")
```

```{r}
#| echo: false
#| results: "asis"
#| label: results

results <- knitr::knit_child(here("scripts/3_results.qmd"), quiet = TRUE)#
cat(results, sep = "\n")
```

```{r}
#| echo: false
#| results: "asis"
#| label: discussion

discussion <- knitr::knit_child(here("scripts/4_discussion.qmd"), quiet = TRUE)
cat(discussion, sep = "\n")
```

```{r}
#| echo: false
#| results: "asis"
#| label: references

references <- knitr::knit_child(here("scripts/5_references.qmd"), quiet = TRUE)
cat(references, sep = "\n")
```

```{r}
#| echo: false
#| results: "asis"
#| label: appendix

appendix <- knitr::knit_child(here("scripts/6_appendix.qmd"), quiet = TRUE)
cat(appendix, sep = "\n")
```

# Statutory Declaration {.unlisted .unnumbered}

```{r declaration}

library(withr)
today <- with_locale(new = c("LC_TIME" = "de_DE"), format(Sys.Date(), "%d. %B %Y"))

```

\thispagestyle{empty}
\newpage

I herewith declare that I have composed the present thesis myself and without use of any other than the
cited sources and aids. Sentences or parts of sentences quoted literally are marked as such; other references
with regard to the statement and scope are indicated by full details of the publications concerned. The thesis
in the same or similar form has not been submitted to any examination body and has not been published.
This thesis was not yet, even in part, used in another examination or as a course performance.

\vspace*{15mm}
\noindent\makebox[3in]{\hrulefill}

\noindent Jan Luca Schnatz

\noindent Darmstadt, den `r today`